Analytic Doppler Imaging

### 2019-08-12 tex/ms.tex
so much text

@@

Next, we expand the spatial dependence of the
specific intensity at the rest frame wavelength
in terms of spherical harmonics on the unit disk:
%
\begin{align}
    \label{eq:deconv:Ixi0}
    I(\lnlam_0, \x, t) 
        &=
        \sum_{l=0}^\infty\sum_{m=-l}^{l} a_{lm}(\lnlam_0, t) Y_{lm}(\x)
    \quad ,
\end{align}
%
where $Y_{lm}(\x)$ is a spherical harmonic on the projected disk
and $a_{lm}(\lnlam_0, t)$ is the corresponding spherical harmonic 
coefficient at log wavelength in the rest frame $\lnlam_0$ and time $t$. For 
convenience, we may write this equation in vector form:
%
\begin{align}
    \label{eq:deconv:Ivec}
    I(\lnlam_0, \x, t) &=
    \ylmbasis(\x) \,
    \alm(\lnlam_0, t)
    \quad ,
\end{align}
%
where
%
\begin{align}
    \label{eq:deconv:alm}
    \alm(\lnlam_0, t) \equiv
\Big( 
    a_{0,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,-1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad
    a_{1,1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)^\top
\end{align}
%
is the vector of spherical harmonic coefficients and
%
\begin{align}
    \ylmbasis(\x) \equiv 
\Big( 
    Y_{0,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,-1}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,1}(\x) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)
\end{align}
%
is the corresponding vector of spherical harmonics. We may further
decompose our expression by linearizing the time dependence of the
spherical harmonic coefficients:
%
\begin{align}
    \label{eq:deconv:R}
    \alm(\lnlam_0, t) = \R(t) \alm(\lnlam_0)
    \quad ,
\end{align}
%
where $\alm(\lnlam_0) = \alm(\lnlam_0, t=t_0)$ for some reference time $t_0$.
For rigid body rotation, the result is exact and $\R(t)$ is the Wigner 
rotation matrix for real spherical harmonics 
\citep[e.g.][]{AlvarezCollado1989}, which is implicitly a 
function of the inclination, obliquity, and rotation period of the star.
In the case that other processes such as differential rotation or spot 
evolution are significant
over the course of an observation, Equation~(\ref{eq:deconv:R}) can be
made to hold approximately for some effective rotation matrix 
$\R(t)$; we discuss this later.

The equation for the specific intensity in the rest frame now reads
%
\begin{align}
    \label{eq:deconv:Ivecfull}
    I(\lnlam_0, \x, t) &=
    \ylmbasis(\x)
    \,
    \R(t)
    \,
    \alm(\lnlam_0)
    \quad ,
\end{align}
%
where it is clear that we have fully separated the spatial, temporal, and
spectral terms. Inserting this into Equation~(\ref{eq:deconv:convolution}) 
and integrating
over the visible portion of the star, we arrive at the expression for the 
observed spectrum:
%
%
\begin{align}
    \label{eq:deconv:F2d}
    F(\lnlam, t, \Dargs) &=
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \,
    \R(t)
    \,
    \alm(\lnlam_0)
    * \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \nonumber \\[0.5em]
    &=
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \,
    \R(t)
    \,
    *
    \,
    \alm(\lnlam_0)
    \quad ,
\end{align}
%
%
where we made use of the commutativity of the convolution operator and 
the fact that the integral is taken only over the spatial dimensions.
Note, importantly, that the convolution operation
above is implicitly a vector operation---i.e., the
convolution is taken for each spherical harmonic term individually
and then summed over all terms.

Finally, we can simplify Equation~(\ref{eq:deconv:F2d}) by noting that the
delta function in the integrand allows us to reduce the double integral 
to a line integral:
%
\begin{align}
    \label{eq:deconv:kT}
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \, \,
    &=  
    \int\limits_{\Curve(\lnlam, \x, \Dargs)}
    \hspace*{-0.6em}\ylmbasis(\x)
    \mathrm{d}\Curve(\lnlam, \x, \Dargs)
    \nonumber \\[0.5em]
    &\equiv \kT(\lnlam, \Dargs)
    \quad.
\end{align}
%
where the path $\Curve(\lnlam, \x, \Dargs)$ corresponds to the
set of all points on the visible disk where 
$\lnlam_0 + \D(\x, \Dargs) = 0$.
%
We thus have
%
\begin{align}
    \label{eq:deconv:F}
    F(\lnlam, t, \Dargs) 
    &=
    \kT(\lnlam, \Dargs) \, \R(t)
    *
    \alm(\lnlam_0)
    \quad.
\end{align}
%
Equation~(\ref{eq:deconv:F}) is the deconvolution of the
observed spectrum into velocity terms, temporal terms, and spectral/spatial
terms, respectively. In the next section, we will discuss how to
efficiently solve the line integral in Equation~(\ref{eq:deconv:kT}).

@@

This completes our derivation of the analytic expression for the
observed spectrum (Equation~\ref{eq:deconv:F}). As we show later,
this equation allows us to compute the forward model for the
spectrum very efficiently, particularly if we make use of Fast-Fourier
Transform (FFT) -based convolution techniques. However, a convolution
can also be represented as a matrix dot product, which we can use to
express the model for the observed spectrum as a purely linear
operation on the map decomposition $\alm(\lnlam_0)$. In the following section
we derive this expression and show how it can be used to obtain
fast, analytic solutions to the Doppler problem.

\subsection{Linearization}
\label{sec:linear}

@@

Given this formulation, we may re-write Equation~(\ref{eq:deconv:F}) as a
purely linear operation on $\alm(\lnlam_0)$:
%
\begin{align}
    \label{eq:linear:ft}
    \bvec{f}_t
    &=
    \Doppler_t
    \,
    \alm
    \quad,
\end{align}
%
where $\Doppler_t$ is a $(K \times N K')$ matrix constructed
by horizontally concatenating the Toeplitz matrices for each of the $N$ 
components of the convolution kernel $\kT(\lnlam, \Dargs)$ after rotating 
them (via $\R(t)$) to the stellar phase at time $t$. 
%
The $(K \times 1)$ vector $\mathbf{f}$ is the model for 
the flux observed at each wavelength at a particular time, and the 
$(N K' \times 1)$ vector $\alm$ is the vector representation of the
spatially-dependent spectrum of the star. The latter is constructed
by flattening the spherical harmonic expansion of the star
such that the first $K'$ terms in $\alm$ correspond
to the the values of $a_{0,0}$ at each wavelength $\lnlam_0$,
followed by the values of $a_{0,0}$ at each wavelength, and so forth.

In general, our data will consist of observations made at several
epochs, corresponding to different rotational phases of the star. 
If we concatenate all $M$ spectra $\bvec{f}_t$ into the $(MK \times 1)$ 
vector $\bvec{f}$, we may write
%
\begin{align}
    \label{eq:linear:f}
    \bvec{f}
    &=
    \Doppler
    \,
    \alm
    \quad,
\end{align}
%
where $\Doppler$ is the full $(MK \times N K')$ Doppler design matrix, 
constructed by vertically concatenating the individual matrices $\Doppler_t$.
This equation represents the full linearization of the problem, where we
have expressed the quantity we can observe, $\bvec{f}$, as a linear
operation on the quantity of interest, the spectral/spatial decomposition
of the stellar surface, $\alm$. Figure~\ref{fig:linalg} shows an
example of $\Doppler$.

\section{The inverse problem}
\label{sec:inverse}
%
In the previous section we showed how to express the data vector $\bvec{f}$,
a series of spectra obtained at different epochs, as a purely linear
operation on the map vector $\alm$, the spherical harmonic decomposition of
the specific intensity on the surface of the star. The advantage of this
linearity is that, given suitable priors, it allows one to compute both 
the maximum likelihood solution for $\alm$ \emph{and} its uncertainty
\emph{analytically}. In particular, if one places a (multidimensional)
Gaussian prior on $\alm$ with mean $\boldsymbol{\mu}$ and 
covariance $\boldsymbol{\Lambda}$, the maximum \emph{a posteriori} (MAP) 
solution for $\alm$ is
%
\begin{align}
    \bvec{\hat{a}} &= 
    \boldsymbol{\Sigma}_\mathbf{\hat{a}}
    \left(
        \Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \bvec{f}
        +
        \boldsymbol{\Lambda}^{-1} \boldsymbol{\mu}
    \right)
    \quad,
\end{align}
%
with posterior covariance given by
%
\begin{align}
    \boldsymbol{\Sigma}_\mathbf{\hat{a}} &= 
    \left(
        \Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \Doppler
        +
        \boldsymbol{\Lambda}^{-1}
    \right)^{-1}
    \quad,
\end{align}
%
where $\boldsymbol{\Sigma}_\mathbf{f}$ is the data covariance
matrix. These equations require the inversion of a few fairly
large matrices, but as we will see later, it allows one to obtain
the map of the star ($\bvec{\hat{a}}$) and its uncertainty
($\boldsymbol{\Sigma}_\mathbf{\hat{a}}$) in under a second for
a typical dataset.

There is, however, a catch: for most practical purposes, it is 
difficult to find a good Gaussian prior for $\alm$. Usually
we will have some prior information on what the spectrum of the
star is, and perhaps some prior information on the distribution
of surface features such as starspots. The problem is that
even if these priors are Gaussian, the corresponding prior
on $\alm$ is \emph{not}. To understand this, consider the
matrix $\bvec{A}$, constructed by reshaping $\alm$ into
a matrix of shape $(N \times K')$, whose rows correspond
to the spectral components associated with each spherical harmonic coefficient.
By performing singular value decomposition (SVD) on this matrix, we may
write
%
\begin{align}
    \bvec{A} = \bvec{U} \, \bvec{S} \, \bvec{V}^\top
\end{align}
%
where $\bvec{U}$ is a matrix of shape $(N \times P)$, $\bvec{V}^\top$ is a 
matrix of shape $(P \times K')$, and $\bvec{S}$ is a diagonal matrix of 
shape $(P \times P)$ for some integer $P$.

Equation~(\ref{eq:deconv:Ivecfull})

\subsection{Continuum normalization}

\subsection{Limb darkening}

@@

\section{Validation}
\label{sec:validation}

Words here. Figure~\ref{fig:spot} and Figure~\ref{fig:compare}.

\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=0.65\linewidth]{figures/compare.pdf}
    \oscaption{compare}{%
        The observed spectrum when the spot is in view computed with
        \starry (blue) and numerically (orange). The problem setup
        is the same as that in Figure~\ref{fig:spot}. The bottom
        panel shows the absolute value of the difference between the 
        two spectra for different grid resolutions in the numerical solution.
        As the number of grid points increases, the numerical solution
        approaches the \starry solution.
        \label{fig:compare}
    }
    \end{centering}
\end{figure}

@@

>>>>

\newcommand{\azero}{\ensuremath{\bvec{a_0}}}

@@

Next, we expand the spatial dependence of the
specific intensity at the rest frame wavelength
in terms of spherical harmonics on the unit disk:
%
\begin{align}
    \label{eq:deconv:Ixi0}
    I(\lnlam_0, \x, t) 
        &=
        \sum_{l=0}^\infty\sum_{m=-l}^{l} a_{lm}(\lnlam_0, t) Y_{lm}(\x)
    \quad ,
\end{align}
%
where $Y_{lm}(\x)$ is a spherical harmonic on the projected disk
and $a_{lm}(\lnlam_0, t)$ is the corresponding spherical harmonic 
coefficient at log wavelength in the rest frame $\lnlam_0$ and time $t$. For 
convenience, we may write this equation in vector form:
%
\begin{align}
    \label{eq:deconv:Ivec}
    I(\lnlam_0, \x, t) &=
    \ylmbasis(\x) \,
    \alm(\lnlam_0, t)
    \quad ,
\end{align}
%
where
%
\begin{align}
    \label{eq:deconv:alm}
    \alm(\lnlam_0, t) \equiv
\Big( 
    a_{0,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,-1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    a_{1,0}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad
    a_{1,1}(\lnlam_0, t) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)^\top
\end{align}
%
is the vector of spherical harmonic coefficients and
%
\begin{align}
    \ylmbasis(\x) \equiv 
\Big( 
    Y_{0,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,-1}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,0}(\x) \quad\quad\quad\quad\quad\quad 
    Y_{1,1}(\x) \quad\quad\quad\quad\quad\quad 
    ... 
\Big)
\end{align}
%
is the corresponding vector of spherical harmonics. We may further
decompose our expression by linearizing the time dependence of the
spherical harmonic coefficients:
%
\begin{align}
    \label{eq:deconv:R}
    \alm(\lnlam_0, t) = \R(t) \, \azero(\lnlam_0)
    \quad ,
\end{align}
%
where $\azero(\lnlam_0) = \bvec{a}(\lnlam_0, t=t_0)$ 
for some reference time $t_0$.
For rigid body rotation, the result is exact and $\R(t)$ is the Wigner 
rotation matrix for real spherical harmonics 
\citep[e.g.][]{AlvarezCollado1989}, which is implicitly a 
function of the inclination, obliquity, and rotation period of the star.
%In the case that other processes such as differential rotation or spot 
%evolution are significant
%over the course of an observation, Equation~(\ref{eq:deconv:R}) can be
%made to hold approximately for some effective rotation matrix 
%$\R(t)$; we discuss this later.

The equation for the specific intensity in the rest frame now reads
%
\begin{align}
    \label{eq:deconv:Ivecfull}
    I(\lnlam_0, \x, t) &=
    \ylmbasis(\x)
    \,
    \R(t)
    \,
    \azero(\lnlam_0)
    \quad ,
\end{align}
%
where it is clear that we have fully separated the spatial, temporal, and
spectral terms. Inserting this into Equation~(\ref{eq:deconv:convolution}) 
and integrating
over the visible portion of the star, we arrive at the expression for the 
observed spectrum:
%
%
\begin{align}
    \label{eq:deconv:F2d}
    F(\lnlam, t, \Dargs) &=
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \,
    \R(t)
    \,
    \alm(\lnlam_0)
    * \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \nonumber \\[0.5em]
    &=
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \,
    \R(t)
    \,
    *
    \,
    \azero(\lnlam_0)
    \quad ,
\end{align}
%
%
where we made use of the commutativity of the convolution operator and 
the fact that the integral is taken only over the spatial dimensions.
Note, importantly, that the convolution operation
above is implicitly a vector operation---i.e., the
convolution is taken for each spherical harmonic term individually
and then summed over all terms.

Finally, we can simplify Equation~(\ref{eq:deconv:F2d}) by noting that the
delta function in the integrand allows us to reduce the double integral 
to a line integral:
%
\begin{align}
    \label{eq:deconv:kT}
    \iint\limits_{\Surf(\x)}
    \ylmbasis(\x)
    \delta\big(\lnlam_0 + \D(\x, \Dargs)\big)
    \mathrm{d}\Surf(\x)
    \, \,
    &=  
    \int\limits_{\Curve(\lnlam, \x, \Dargs)}
    \hspace*{-0.6em}\ylmbasis(\x)
    \mathrm{d}\Curve(\lnlam, \x, \Dargs)
    \nonumber \\[0.5em]
    &\equiv \kT(\lnlam, \Dargs)
    \quad.
\end{align}
%
where the path $\Curve(\lnlam, \x, \Dargs)$ corresponds to the
set of all points on the visible disk where 
$\lnlam_0 + \D(\x, \Dargs) = 0$.
%
We thus have
%
\begin{align}
    \label{eq:deconv:F}
    F(\lnlam, t, \Dargs) 
    &=
    \kT(\lnlam, \Dargs) \, \R(t)
    *
    \azero(\lnlam_0)
    \quad.
\end{align}
%
Equation~(\ref{eq:deconv:F}) is the deconvolution of the
observed spectrum into velocity terms, temporal terms, and spectral/spatial
terms, respectively. In the next section, we will discuss how to
efficiently solve the line integral in Equation~(\ref{eq:deconv:kT}).

@@

This completes our derivation of the analytic expression for the
observed spectrum (Equation~\ref{eq:deconv:F}).

\subsection{Example}
%
Figure~\ref{fig:spot} shows an example application of the formulae derived 
in this section. We used \starry to generate a synthetic stellar surface with a 
single large Gaussian spot at a latitude of $30^\circ$, yielding a vector of
spherical harmonic coefficients $\bvec{y}$.
The stellar rest frame spectrum $I(\lnlam_0)$ is taken to be a single Gaussian 
absorption line that is spatially constant 
everywhere save for a multiplicative factor equal to the local spot intensity.
This corresponds to a spectral/spatial expansion vector $\azero(\lnlam_0)$ 
whose coefficient at degree $l$ and order $m$ is equal to $y_{lm} I(\lnlam_0)$.
In the Figure, we rotated the star about an axis perpendicular to the line
of sight (left panel) and computed the observed spectrum using 
Equation~(\ref{eq:deconv:F}) (center panel). The right panel shows the difference
between the observed spectrum and the spectrum when the spot is not in view.
The effect of the spot on the absorption line is evident: a ``bump'' that
travels from the blueshifted side of to the redshifted side of the line
in sync with the change in longitude of the spot. This correspondence between
line shape residuals and spot location is the cornerstone of the Doppler
imaging technique \citep[compare to, e.g., Figures 1 and 4 in][]{Vogt1983}.

Figure~\ref{fig:compare} shows one of the spectra in Figure~\ref{fig:spot},
this time computed using both our formulae (blue line) and the traditional
technique of discretizing the stellar surface, Doppler-shifting the spectrum
in each grid cell according to the local radial velocity, interpolating back
to a uniform wavelength grid, and summing over the spatial dimension
(dotted orange line). We show the residuals in the lower panel, where it
is clear that as the number of grid cells increases from 2,000 (light grey)
to 71,000 (dark grey), the numerical solution approaches the solution
obtained using our approach. \xxx{Understand---and comment on---the fundamental
differences between a discrete linear convolution and traditional linear 
interpolation. Should we expect the numerical solution to actually \emph{converge}
to the \starry solution, or is the way the two are interpolating fundamentally
different?}

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=0.65\linewidth]{figures/compare.pdf}
    \oscaption{compare}{%
        The observed spectrum when the spot is in view computed with
        \starry (blue) and numerically (orange). The problem setup
        is the same as that in Figure~\ref{fig:spot}. The bottom
        panel shows the absolute value of the difference between the 
        two spectra for different number of grid points in the 
        numerical solution.
        As the number of grid points increases, the numerical solution
        approaches the \starry solution.
        \label{fig:compare}
    }
    \end{centering}
\end{figure}

\section{Linearization}
\label{sec:linear}

@@

Given this formulation, we may re-write Equation~(\ref{eq:deconv:F}) as a
purely linear operation on $\azero(\lnlam_0)$:
%
\begin{align}
    \label{eq:linear:ft}
    \bvec{f}_t
    &=
    \Doppler_t
    \,
    \azero
    \quad,
\end{align}
%
where $\Doppler_t$ is a $(K \times N K')$ matrix constructed
by horizontally concatenating the Toeplitz matrices for each of the $N$ 
components of the convolution kernel $\kT(\lnlam, \Dargs)$ after rotating 
them (via $\R(t)$) to the stellar phase at time $t$. 
%
The $(K \times 1)$ vector $\mathbf{f}$ is the model for 
the flux observed at each wavelength at a particular time, and the 
$(N K' \times 1)$ vector $\azero$ is the vector representation of the
spatially-dependent spectrum of the star. The latter is constructed
by flattening the spherical harmonic expansion of the star
such that the first $K'$ terms in $\azero$ correspond
to the the values of $(a_0)_{0,0}$ at each wavelength $\lnlam_0$,
followed by the values of $(a_0)_{1,-1}$ at each wavelength, and so forth.

In general, our data will consist of observations made at several
epochs, corresponding to different rotational phases of the star. 
If we concatenate all $M$ spectra $\bvec{f}_t$ into the $(MK \times 1)$ 
vector $\bvec{f}$, we may write
%
\begin{align}
    \label{eq:linear:f}
    \bvec{f}
    &=
    \Doppler
    \,
    \azero
    \quad,
\end{align}
%
where $\Doppler$ is the full $(MK \times N K')$ Doppler design matrix, 
constructed by vertically concatenating the individual matrices $\Doppler_t$.
This equation represents the full linearization of the problem, where we
have expressed the quantity we can observe, $\bvec{f}$, as a linear
operation on the quantity of interest, the spectral/spatial decomposition
of the stellar surface, $\azero$. Figure~\ref{fig:linalg} shows an
example of $\Doppler$.

\section{The inverse problem}
\label{sec:inverse}
%
In the previous section we showed how to express the data vector $\bvec{f}$,
a series of spectra obtained at different epochs, as a purely linear
operation on the map vector $\azero$, the spherical harmonic decomposition of
the specific intensity on the surface of the star. The advantage of this
linearity is that, given suitable priors, it allows one to compute both 
the maximum likelihood solution for $\azero$ and its uncertainty
\emph{analytically}. In particular, if one places a (multidimensional)
Gaussian prior on $\azero$ with mean $\boldsymbol{\mu}_\mathbf{a_0}$ and 
covariance $\boldsymbol{\Lambda}_\mathbf{a_0}$, the maximum \emph{a posteriori} (MAP) 
solution for $\azero$ is
%
\begin{align}
    \label{eq:inverse:ahat}
    \bvec{\hat{a}_0} &= 
    \boldsymbol{\Sigma}_\mathbf{\hat{a}_0}
    \left(
        \Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \bvec{f}
        +
        {\boldsymbol{\Lambda}_{\azero}}^{-1} \boldsymbol{\mu}_\mathbf{a_0}
    \right)
    \quad,
\end{align}
%
with posterior covariance given by
%
\begin{align}
    \label{eq:inverse:acov}
    \boldsymbol{\Sigma}_\mathbf{\hat{a}_0} &= 
    \left(
        \Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \Doppler
        +
        {\boldsymbol{\Lambda}_{\azero}}^{-1}
    \right)^{-1}
    \quad,
\end{align}
%
where $\boldsymbol{\Sigma}_\mathbf{f}$ is the data covariance
matrix. These equations require the inversion of a few fairly
large matrices, but as we will see later, it allows one to obtain
the map of the star ($\bvec{\hat{a}_0}$) and its uncertainty
($\boldsymbol{\Sigma}_{\mathbf{\hat{a}_0}}$) in under a second for
a typical dataset.

There is, however, a catch: for most practical purposes, it is 
difficult to find a good Gaussian prior for $\bvec{a_0}$. Usually
we will have some prior information on what the spectrum of the
star is, and perhaps some prior information on the distribution
of surface features such as starspots. The problem is that
even if these priors are Gaussian, the corresponding prior
on $\bvec{a_0}$ is \emph{not}. To understand this, consider the
case where the stellar rest frame spectrum is spatially constant,
save for an amplitude that is spatially variable. We can then
describe the rest frame spectrum by the $(K' \times 1)$ vector
$\bvec{s}$ and the amplitude by the $(N \times 1)$ vector 
$\bvec{y}$ representing the spherical harmonic expansion of the
intensity profile of the star. The vector $\azero$ is then
given by the (flattened) outer product of the two vectors:
%
\begin{align}
    \azero &= \mathrm{vec}\left( \bvec{A_0} \right) \nonumber \\
         &= \mathrm{vec}\left( \bvec{s} \, \bvec{y}^\top \right) \quad,
\end{align}
%
where $\mathrm{vec}$ denotes the vectorization operation, which in this case
transforms the $(K' \times N)$ matrix $\bvec{A_0}$ into the $(N K' \times 1)$
column vector $\azero$ by vertically stacking all of its columns. In other
words, the vector $\azero$ consists of the set of $K'$ intensities in
each wavelength bin repeated $N$ times, each time multiplied by the 
$n^\mathrm{th}$ spherical harmonic coefficient in the expansion of the 
surface intensity. Returning to our point about priors, note that all of the
entries of $\azero$ are the product of two independent random variables. Since
the product of two random variables is generally non-Gaussian, even
when the variables themselves are Gaussian%
\footnote{see, e.g., \url{http://mathworld.wolfram.com/NormalProductDistribution.html}}%
, so too is the prior on $\azero$. This point makes
Equation~(\ref{eq:inverse:ahat}) of little practical use, since a 
multivariate Gaussian will not 
generally be a suitable prior for $\azero$.

That said, a Gaussian prior \emph{is} typically suitable for both the spectrum
$\bvec{s}$ and the intensity $\bvec{y}$ (just not their product). It is
straightforward to show that because Equation~(\ref{eq:linear:f}) is linear
in $\azero$, it is also linear in both $\bvec{s}$ and $\bvec{y}$, so
MAP solutions can be found analytically for $\bvec{s}$ 
(conditioned on $\bvec{y}$) and $\bvec{y}$ (conditioned on $\bvec{s}$). We
investigate these solutions below.

\subsection{Solution for the intensity}
For a fixed, spatially constant rest-frame spectrum $\bvec{s}$, the model
for the flux vector may be written as
%
\begin{align}
    \label{eq:inverse:fy}
    \bvec{f}
    &=
    \Doppler
    \,
    \bvec{S}
    \,
    \bvec{y}
    \quad,
\end{align}
%
where
%
\begin{align}
    \setstackgap{L}{1.25\baselineskip}
    \bvec{S} =
        \begin{pmatrix}
        \quadquad\bvec{s}\quadquad & &            &            &  \\
        & \quadquad\bvec{s}\quadquad &            &            &  \\
        &            & \quadquad\bvec{s}\quadquad &            &  \\
        &            &            & \quadquad\bvec{s}\quadquad &  \\
        &            &            &               & \ddots
        \end{pmatrix}
\end{align}
%
is a $(NK' \times N)$ block matrix constructed by repeating the column
vector $\bvec{s}$ $N$ times in blocks along the main diagonal. Given
prior mean $\boldsymbol{\mu}_\mathbf{y}$ and covariance
$\boldsymbol{\Lambda}_\mathbf{y}$ on $\mathbf{y}$, the MAP solution
is given by
%
\begin{align}
    \label{eq:inverse:yhat}
    \bvec{\hat{y}} &= 
    \boldsymbol{\Sigma}_\mathbf{\hat{y}}
    \left(
        \bvec{S}^\top\Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \bvec{f}
        +
        {\boldsymbol{\Lambda}_\mathbf{y}}^{-1} \boldsymbol{\mu}_\mathbf{y}
    \right)
    \quad,
\end{align}
%
with covariance
%
\begin{align}
    \label{eq:inverse:ycov}
    \boldsymbol{\Sigma}_\mathbf{\hat{y}} &= 
    \left(
        \bvec{S}^\top\Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \Doppler\bvec{S}
        +
        {\boldsymbol{\Lambda}_\mathbf{y}}^{-1}
    \right)^{-1}
    \quad.
\end{align}
%
If we choose a prior with zero mean ($\boldsymbol{\mu}_\bvec{y} = 0$)
and a diagonal covariance that is constant for each spherical harmonic 
order $m$,
%
\begin{align}
    \boldsymbol{\Lambda}_\bvec{y} &=
    \mathrm{diag} \left(
        \sigma_0^2
        \quad\quad\quad\quad
        \sigma_1^2
        \quad\quad\quad\quad
        \sigma_1^2
        \quad\quad\quad\quad
        \sigma_1^2
        \quad\quad\quad\quad
        \sigma_2^2
        \quad\quad\quad\quad
        \sigma_2^2
        \quad\quad\quad\quad
        \sigma_2^2
        \quad\quad\quad\quad
        \sigma_2^2
        \quad\quad\quad\quad
        \sigma_2^2
        \quad\quad\quad\quad
        \cdots
    \right)
    \quad,
\end{align}
%
we are in fact imposing an isotropic prior with an angular power spectrum
given by $C_l = \sigma_l^2$
\citep[e.g.,][]{Baldi2006}. This is especially useful if the typical
angular scale of features on the surface of the star is known or if
a Gaussian prior can be placed on it.

\subsection{Solution for the spectrum}
If, on the other hand, the spatial intensity profile is known, the
model for the flux vector may be written in terms of the spectrum $\bvec{s}$
as
%
\begin{align}
    \label{eq:inverse:fs}
    \bvec{f}
    &=
    \Doppler
    \,
    \bvec{Y}
    \,
    \bvec{s}
    \quad,
\end{align}
%
where
%
\begin{align}
    \setstackgap{L}{1.25\baselineskip}
    \bvec{Y} =
        \begin{pmatrix}
            y_0\, \bvec{I}_{K'} \\
            y_1\, \bvec{I}_{K'} \\
            y_2\, \bvec{I}_{K'} \\
            y_3\, \bvec{I}_{K'} \\
            \cdots
        \end{pmatrix}
\end{align}
%
is a $(NK' \times K')$ matrix constructed by vertically stacking
$N$ $(K' \times K')$ identity matrices, each multiplied by the
corresponding coefficient of $\bvec{y}$. Given
prior mean $\boldsymbol{\mu}_\mathbf{s}$ and covariance
$\boldsymbol{\Lambda}_\mathbf{s}$ on $\mathbf{s}$, the MAP solution
is given by
%
\begin{align}
    \label{eq:inverse:shat}
    \bvec{\hat{s}} &= 
    \boldsymbol{\Sigma}_\mathbf{\hat{s}}
    \left(
        \bvec{Y}^\top\Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \bvec{f}
        +
        {\boldsymbol{\Lambda}_\mathbf{s}}^{-1} \boldsymbol{\mu}_\mathbf{s}
    \right)
    \quad,
\end{align}
%
with covariance
%
\begin{align}
    \label{eq:inverse:scov}
    \boldsymbol{\Sigma}_\mathbf{\hat{s}} &= 
    \left(
        \bvec{Y}^\top\Doppler^\top
        {\boldsymbol{\Sigma}_\mathbf{f}}^{-1}
        \Doppler\bvec{Y}
        +
        {\boldsymbol{\Lambda}_\mathbf{s}}^{-1}
    \right)^{-1}
    \quad.
\end{align}

\subsection{The full solution}
In the previous two sections we showed how, if the stellar
rest-frame spectrum is spatially constant and known \emph{a priori} 
(from, say, a template spectrum), the posterior distribution
for the surface map $\bvec{y}$ is \emph{analytic} 
(Equations~\ref{eq:inverse:yhat}
and \ref{eq:inverse:ycov}). Conversely, if the surface map is known,
the posterior for the spectrum is also analytic 
(Equations~\ref{eq:inverse:shat} and \ref{eq:inverse:scov}). Both sets of
solutions require Gaussian priors on the quantities of interest, but these
are easily interpretable. A Gaussian prior on
the spherical harmonic coefficients of the surface map corresponds 
to the angular power spectrum of the surface features, whereas a Gaussian
prior on the spectrum can be used to correctly propagate uncertainties in
the stellar spectrum from observational uncertainties on the stellar
metallicity, temperature, etc.

Importantly, the framework developed here allows one to analytically
compute not only the ``optimal'' solution to the Doppler imaging problem,
but also its uncertainty. The problem of re-constructing surface images
of stars from spectra and/or light curves is fundamentally ill-posed because
of intrinsic degeneracies and a large null space
\citep[e.g.,][]{Luger2019} \xxx{more cites}. Coupled to uncertainties in
the observations themselves, this makes the process of searching for the
``optimal'' solution flawed at best, and meaningless at worst. Access to the
full covariance matrix of the solution is therefore essential to assessing our
confidence in the various features of the inferred surface maps.

However, as we mentioned above, the Doppler imaging problem can be solved
analytically for either the surface map or the spectrum, but not both. If
neither are known exactly \emph{a priori}, which is quite often the case,
we can still find the MAP solution quickly by guessing at an initial
spectrum $\bvec{s}$ and iterating between optimizing for $\bvec{y}$ and
optimizing for $\bvec{s}$. This bi-linear problem is not generally convex,
but given a good starting point for $\bvec{s}$ and sufficient regularization
on $\bvec{y}$, we find that it typically converges to the correct solution,
especially after refinement with a non-linear solver;
we discuss this in detail below.

Once the MAP solution has been found for both $\bvec{s}$ and $\bvec{y}$,
the covariance of one conditioned on the other can be computed from the
formulae above. However, this is necessarily a lower bound on the true
covariance of the solution, since it neglects any covariance \emph{between}
$\bvec{s}$ and $\bvec{y}$. To \emph{correctly} account for this, one must
resort to Monte Carlo methods or other sampling techniques. We discuss
this further below.

@@

@@

\cite{Vogt1987}
